---
title: "Example"
author: "Ross Cunning"
date: "2025-04-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r libraries}
# Load libraries
library(tidyverse)
library(ggpubr)
```

```{r functions}
# Define helper functions

# Function to test if slopes are significantly different from null slope
slope_test <- function(model, null_slope = 0, alternative = c("less", "greater", "two.sided")) {
  alternative <- match.arg(alternative)

  # Extract slope estimate and SE
  coefs <- summary(model)$coefficients
  slope_est <- coefs[2, "Estimate"]
  slope_se  <- coefs[2, "Std. Error"]
  df_resid  <- model$df.residual

  # Compute t-statistic
  t_stat <- (slope_est - null_slope) / slope_se

  # Compute p-value
  p_val <- switch(alternative,
                  less = pt(t_stat, df = df_resid),
                  greater = 1 - pt(t_stat, df = df_resid),
                  two.sided = 2 * pt(-abs(t_stat), df = df_resid))

  # Map alternative to null hypothesis symbol
  null_symbol <- switch(alternative,
                        less = "≥",
                        greater = "≤",
                        two.sided = "=")

  # Get predictor and response variable names
  terms <- all.vars(formula(model))
  xvar <- terms[2]
  yvar <- terms[1]
  data <- model$model

  # Create ggplot layers
  line_layer <- geom_smooth(method = "lm", se = FALSE, color = "blue")
  text_layer <- annotate("text", color = "blue",
                         x = min(data[[xvar]], na.rm = TRUE),
                         y = max(data[[yvar]], na.rm = TRUE),
                         label = paste0("Slope = ", round(slope_est, 3),
                                        "\nP (H0: slope ", null_symbol, " ", null_slope, ") = ",
                                        signif(p_val, 3)),
                         hjust = 0, vjust = 1)

  list(line = line_layer, annotation = text_layer)
}



# Function to correct for spurious slopes, from Kelly and Price 2005
rttm.adj <- function(m1, m2){
  raw.growth<-m2-m1
  vart<-var.test(m1,m2,paired = T) ## variances equal? 
  vpv<-vart$p.value # var.test p value
  m1m2cor<-cor.test(m1, m2) # test correlation between m1 and m2 
  rho<-m1m2cor$estimate # correlation coefficient between m1 and m2 
  m1sd<-sd(m1) # m1 sd
  m2sd<-sd(m2) # m2 sd
  m1v<-var(m1) # m1 var
  m2v<-var(m2) # m2 var
  m1m<-mean(m1) # m1 mean
  m2m<-mean(m2) # m2 mean
  pm<-mean(raw.growth)
  rho2<-(2*rho*m1sd*m2sd)/(m1v+m2v) # adjusted correlation coefficient used if variances are equal
  rhof<-ifelse(vpv <= 0.05, rho, rho2) # which rho is used for dstar calculation is based on variance comparison
  dstar<-(rhof*(m1-m1m)-(m2-m2m))*-1 # adjustment values. Multiply by -1 to flip sign because Kelly and Price based on plasticity as m1-m2, not m2-m1 as in most thermal tolerance estimates
  adj.growth <- pm+dstar # corrected plasticity. 
  out<-as.data.frame(cbind(raw.growth, dstar, adj.growth)) 
  return(out)
}
```

# Set parameters for simulations
```{r}
# Get mean and standard deviations from real data
pdam0 <- read_csv("PdamRbleaching.csv")
pdam_summ <- pdam0 %>% 
  group_by(sym) %>%
  summarise(across(c(juntotal, augtotal), 
                   list(mean = ~mean(log(.)), sd = ~sd(log(.)))))

# Get mean initial S/H in each group corals (jun)
pdam.mean.init <- mean(pdam_summ$juntotal_mean)
# Get mean final S/H in each group of corals (aug)
pdam.mean.final <- mean(pdam_summ$augtotal_mean)

# Get average SD of S/H ratio in each group of corals (symbiont / timepoint)
pdam.sd <- mean(c(pdam_summ$juntotal_sd, pdam_summ$augtotal_sd))

# Get mean and SD of log changes in S/H ratio across all corals
pdam.mean.logchange <- mean(log(pdam0$augtotal/pdam0$juntotal))
pdam.sd.logchange <- sd(log(pdam0$augtotal/pdam0$juntotal))
  

# Set number of individuals for simulations
n <- 10000
```

# Scenario 1: Relative change is unrelated to initial value; true values known (no measurement error)
```{r s1}
# Goal: Confirm that if log_change is independent of log_initial, regression shows no spurious slope
# assumes log percent of symbionts lost is random normal

set.seed(1)
df1 <- tibble(
  log_init_true = rnorm(n, mean = pdam.mean.init, sd = pdam.sd),    # Initial values (log)
  log_change_true = log(0.25) + rnorm(n, mean = 0, sd = 0.3),       # Independent log change, 75% [log(final/initial)] + noise
  log_final_true = log_init_true + log_change_true                  # Final = initial + change
)

# Final vs. initial, true values
# Expect: slope ≈ 1
mod1.1 <- lm(log_final_true ~ log_init_true, data = df1)
res1.1 <- slope_test(mod1.1, 1, "less")
p1.1 <- ggplot(df1, aes(x = log_init_true, y = log_final_true)) +
  geom_point(alpha = 0.25) +
  res1.1$line +
  res1.1$annotation +
  theme_minimal() +
  labs(title = "Final vs. initial (true values)")

# Change vs. initial, true values
# Expect: slope ≈ 0
mod1.2 <- lm(log_change_true ~ log_init_true, data = df1)
res1.2 <- slope_test(mod1.2, 0, "less")
p1.2 <- ggplot(df1, aes(x = log_init_true, y = log_change_true)) +
  geom_point(alpha = 0.25) +
  res1.2$line +
  res1.2$annotation +
  theme_minimal() +
  labs(title = "Change vs. initial (no relationship; true values)")

p1.2
```

Confirms that there is no spurious relationship in change vs. initial values


# Scenario 2: Relative change is unrelated to initial value; values are measured with error
```{r s2.1}
set.seed(1)
df2 <- df1 %>%
  mutate(log_init_obs = log_init_true + rnorm(n, mean = 0, sd = 0.3),   # Add random measurement error to initial values
         log_final_obs = log_final_true + rnorm(n, mean = 0, sd = 0.3), # Add random measurement error to final values
         log_change_obs = log_final_obs - log_init_obs)                 # Observed change

# Final vs. Initial, observed values with measurement error
# Expect: slope ≈ 1
mod2.1 <- lm(log_final_obs ~ log_init_obs, data = df2)
res2.1 <- slope_test(mod2.1, 1, "less")
p2.1 <- ggplot(df2, aes(x = log_init_obs, y = log_final_obs)) +
  geom_point(alpha = 0.25) +
  res2.1$line +
  res2.1$annotation +
  theme_minimal() +
  labs(title = "Final vs. initial (no relationship; with measurement error)")

# Change vs. initial, observed values with measurement error
# Expect: slope ≈ 0
mod2.2 <- lm(log_change_obs ~ log_init_obs, data = df2)
res2.2 <- slope_test(mod2.2, 0, "less")
p2.2 <- ggplot(df2, aes(x = log_init_obs, y = log_change_obs)) +
  geom_point(alpha = 0.25) +
  res2.2$line +
  res2.2$annotation +
  theme_minimal() +
  labs(title = "Change vs. initial (no relationship; with measurement error)")

p2.2
```

Shows measurement error in initial values leads to mathematical coupling and spurious relationship

### Does the Kelly and Price method recover correct relationship (slope = 0)?
```{r s2.2}
# Apply Kelly and Price method to get adjusted change
df2 <- df2 %>%
  mutate(log_change_obs_adj = rttm.adj(log_init_obs, log_final_obs)[, 3])

# Model adjusted change vs. initial values
mod2.3 <- lm(log_change_obs_adj ~ log_init_obs, data = df2)
res2.3 <- slope_test(mod2.3, 0, "less")
p2.3 <- ggplot(df2, aes(x = log_init_obs, y = log_change_obs_adj)) +
  geom_point(alpha = 0.25) +
  res2.3$line +
  res2.3$annotation +
  theme_minimal() +
  labs(title = "K-P adjusted change vs. initial (no relationship; with measurement error)")

p2.3
```

Yes, Kelly and Price removes the spurious relationship caused by mathematical coupling. It is also able to do this when the measurement error is more extreme (tested with sd = 1).


# Scenario 3: True negative relationship between change and initial values
```{r 3.1}
# Goal: Simulate a case where more symbionts causes greater loss (true relationship)

# Define strength of initial value effect on logchange
# e.g., -0.2 means for every 1 unit increase in initial, 0.2 unit decrease in logchange
b <- -0.5

# Define intercept to target 75% loss at mean initial value
intercept <- log(0.25) - b * mean(df1$log_init_true)

set.seed(6)
# Simulate log changes as an intercept + effect of initial value + noise
df3 <- df1 %>%
  mutate(log_change_true = intercept + b * log_init_true + rnorm(n, mean = 0, sd = 0.3),
         log_final_true = log_init_true + log_change_true)

cor(df3$log_init_true, df3$log_change_true)

# Final vs. initial, true values
# Expected: slope = 0.8
mod3.1 <- lm(log_final_true ~ log_init_true, data = df3)
res3.1 <- slope_test(mod3.1, 1, "less")
p3.1 <- ggplot(df3, aes(x = log_init_true, y = log_final_true)) +
  geom_point(alpha = 0.25) +
  res3.1$line +
  res3.1$annotation +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  theme_minimal() +
  labs(title = "Final vs. initial (true relationship; true values)")

# Change vs. initial, true values
# Expected: slope = -0.2
mod3.2 <- lm(log_change_true ~ log_init_true, data = df3)
res3.2 <- slope_test(mod3.2, 0, "less")
p3.2 <- ggplot(df3, aes(x = log_init_true, y = log_change_true)) +
  geom_point(alpha = 0.25) +
  res3.2$line +
  res3.2$annotation +
  theme_minimal() +
  labs(title = "Change vs. initial (true relationship; true values)")

p3.2
```

### Does the Kelly and Price method change this relationship?
```{r s3.2}
# Apply Kelly and Price method to get adjusted change
df3 <- df3 %>%
  mutate(log_change_true_adj = rttm.adj(log_init_true, log_final_true)[, 3])

# Model adjusted change vs. initial values
mod3.3 <- lm(log_change_true_adj ~ log_init_true, data = df3)
res3.3 <- slope_test(mod3.3, 0, "less")
p3.3 <- ggplot(df3, aes(x = log_init_true, y = log_change_true_adj)) +
  geom_point(alpha = 0.25) +
  res3.3$line +
  res3.3$annotation +
  theme_minimal() +
  labs(title = "K-P adjusted change vs. initial (true relationship; true values)")

p3.3
```

Need to test this more, but it seems to reduce the relationship a lot, and all the way to zero for lower values of b. Potentially there's a threshold b effect size wheren the adjustment still produces a negative but diminished relationship

Does the measurement error part come into effect here???

# Scenario 4: Real data
```{r s4}
pdam <- pdam0 %>%
  mutate(log_init_raw = log(juntotal),
         log_final_raw = log(augtotal),
         log_change = log_final_raw - log_init_raw)

# ggplot(pdam, aes(x = log_init_raw, y = log_change, shape = sym, color = sym)) +
#   geom_point() +
#   geom_smooth(method = "lm", se = FALSE) +
#   theme_minimal()

# Remove symbiont effect
# Calculate grand mean
grand_mean_init <- mean(pdam$log_init_raw)
grand_mean_final <- mean(pdam$log_final_raw)
# Add sym-group residuals to grand mean
pdam <- pdam %>%
  group_by(sym) %>%
  mutate(group_resid_init = log_init_raw - mean(log_init_raw),
         group_resid_final = log_final_raw - mean(log_final_raw)) %>%
  ungroup() %>%
  mutate(log_init = grand_mean_init + group_resid_init,
         log_final = grand_mean_final + group_resid_final) %>%
  select(colony, sym, log_init_raw, log_final_raw, log_init, log_final) %>%
  mutate(log_change = log_final - log_init)

# ggplot(pdam, aes(x = log_init, y = log_change, shape = sym, color = sym)) +
#   geom_point() +
#   geom_smooth(method = "lm", se = FALSE) +
#   theme_minimal()

# Change vs. initial, real data
# Expected: slope = -0.2
mod4.2 <- lm(log_change ~ log_init, data = pdam)
res4.2 <- slope_test(mod4.2, 0, "less")
p4.2 <- ggplot(pdam, aes(x = log_init, y = log_change)) +
  geom_point(alpha = 0.25) +
  res4.2$line +
  res4.2$annotation +
  theme_minimal() +
  labs(title = "Change vs. initial (real data)")

p4.2
```

### Does the Kelly and Price method change this relationship?
```{r s3.2}
# Test for bivariate normality
MVN::mvn(select(pdam, log_init, log_final))

# Apply Kelly and Price method to get adjusted change
pdam <- pdam %>%
  mutate(log_change_adj = rttm.adj(log_init, log_final)[, 3])

# Model adjusted change vs. initial values
mod4.3 <- lm(log_change_adj ~ log_init, data = pdam)
res4.3 <- slope_test(mod4.3, 0, "less")
p4.3 <- ggplot(pdam, aes(x = log_init, y = log_change_adj)) +
  geom_point(alpha = 0.25) +
  res4.3$line +
  res4.3$annotation +
  theme_minimal() +
  labs(title = "K-P adjusted change vs. initial (real data)")

p4.3

# back-transform adjusted data?
# pdam <- pdam %>%
#   mutate(back_change = exp(log_change_adj),
#          back_init = exp(log_init))
# ggplot(pdam, aes(x = back_init, y = back_change)) + geom_point()
```

# Real data no log?
```{r}
# Try on S/H ratios (non-logged)
pdam <- pdam %>%
  mutate(init = exp(log_init),
         final = exp(log_final),
         change = final - init)
MVN::mvn(select(pdam, init, final))     # Does not meet bivariate normal assumption
  
# Plot
ggplot(pdam, aes(x = init, y = change)) + geom_point()
summary(lm(change~init, data = pdam))

# Adjustment
pdam <- pdam %>%
  mutate(change_adj = rttm.adj(init, final)[, 3])

# Replot
ggplot(pdam, aes(x = init, y = change_adj)) + geom_point()
summary(lm(change_adj~init, data = pdam))
```

# Real data Million Acer growth
```{r}
mill <- tibble(subMil1) %>%
  filter(size > 0) %>%  # there are four rows where size = 0, filter out because will generate Inf
  filter(InitialSize > 2) %>%     # cutoff of 1 gives adjusted marginal sig, cutoff 2 nonsig
  mutate(init = InitialSize,
         final = size,
         change = final - init)
MVN::mvn(data = select(mill, init, final))     # NOT multivariate normal, violates Kelly and Price assumption...

# Change vs. initial
mod5.2 <- lm(change ~ init, data = mill)
res5.2 <- slope_test(mod5.2, 0, "greater")
p5.2 <- ggplot(mill, aes(x = init, y = change)) +
  geom_point(alpha = 0.25) +
  res5.2$line +
  res5.2$annotation +
  theme_minimal() +
  labs(title = "Change vs. initial")

p5.2

# Adjusted change
mill <- mill %>%
  mutate(change_adj = rttm.adj(init, final)[, 3])

mod5.3 <- lm(change_adj ~ init, data = mill)
res5.3 <- slope_test(mod5.3, 0, "greater")
p5.3 <- ggplot(mill, aes(x = init, y = change_adj)) +
  geom_point(alpha = 0.25) +
  res5.3$line +
  res5.3$annotation +
  theme_minimal() +
  labs(title = "Adjusted Change vs. initial")

p5.3
```

This reproduces Carly's analysis --- but problem of lognormal?

# log-transform size
```{r}
mill <- tibble(subMil1) %>%
  filter(size > 0) %>%  # there are four rows where size = 0, filter out because will generate Inf
  filter(InitialSize > 2) %>%     # cutoff of 1 gives adjusted marginal sig, cutoff 2 nonsig
  mutate(log_init = log(InitialSize),
         log_final = log(size),
         log_change = log_final - log_init) %>%
  filter(log_change > -1 & log_change < 2)

# Change vs. initial
mod5.2 <- lm(log_change ~ log_init, data = mill)
res5.2 <- slope_test(mod5.2, 0, "less")
p5.2 <- ggplot(mill, aes(x = log_init, y = log_change)) +
  geom_point(alpha = 0.25) +
  res5.2$line +
  res5.2$annotation +
  theme_minimal() +
  labs(title = "Change vs. initial")

p5.2

# Adjusted change
mill <- mill %>%
  mutate(log_change_adj = rttm.adj(log_init, log_final)[, 3])

mod5.3 <- lm(log_change_adj ~ log_init, data = mill)
res5.3 <- slope_test(mod5.3, 0, "less")
p5.3 <- ggplot(mill, aes(x = log_init, y = log_change_adj)) +
  geom_point(alpha = 0.25) +
  res5.3$line +
  res5.3$annotation +
  theme_minimal() +
  labs(title = "Adjusted Change vs. initial")

p5.3

# backtransform adjusted sizes
mill <- mill %>%
  mutate(back_change_adj = exp(log_change_adj))
ggplot(mill, aes(x = InitialSize, y = back_change_adj)) + geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```



