---
title: "Figure 3 alternatives for reviewer response"
date: "May 2026"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE)
```

# Purpose

Figure 3 mirrors Figure 1 for the two empirical case studies. This notebook
creates the same family of alternatives used for Figure 1:

1. original recreated;
2. original structure with adjusted baseline-change panels removed;
3. reviewer-minimal;
4. compact overlay.

# Setup

```{r libraries}
library(tidyverse)
library(cowplot)
library(ggpubr)
library(boot)
```

```{r helpers}
dir.create("figures/figure3_alternatives", recursive = TRUE, showWarnings = FALSE)

method_colors <- c(
  "Observed" = "#0072B2",
  "Kelly-Price" = "#D55E00",
  "Blomqvist" = "#009E73",
  "Moments" = "#CC79A7",
  "Obs." = "#0072B2",
  "K.-P." = "#D55E00",
  "Blom." = "#009E73",
  "Mom." = "#CC79A7"
)

method_linetypes <- c(
  "Observed" = "solid",
  "Kelly-Price" = "22",
  "Blomqvist" = "42",
  "Moments" = "73"
)

format_p_label <- function(p) {
  if (is.na(p)) return("p = NA")
  if (p < 2.2e-16) return("p < 2.2e-16")
  if (p < 0.001) return(paste0("p = ", formatC(p, format = "e", digits = 1)))
  paste0("p = ", formatC(p, format = "fg", digits = 3))
}

rttm.adj <- function(m1, m2) {
  raw.growth <- m2 - m1
  vart <- var.test(m1, m2, paired = TRUE)
  vpv <- vart$p.value
  rho <- cor.test(m1, m2)$estimate
  m1sd <- sd(m1)
  m2sd <- sd(m2)
  m1v <- var(m1)
  m2v <- var(m2)
  m1m <- mean(m1)
  m2m <- mean(m2)
  pm <- mean(raw.growth)
  rho2 <- (2 * rho * m1sd * m2sd) / (m1v + m2v)
  rhof <- ifelse(vpv <= 0.05, rho, rho2)
  dstar <- (rhof * (m1 - m1m) - (m2 - m2m)) * -1
  data.frame(raw.growth = raw.growth, dstar = dstar, adj.growth = pm + dstar)
}

blomqvist_b <- function(b_obs, k) {
  (b_obs + k) / (1 - k)
}

blomqvist_pseudo_exact <- function(x, y, b_blom) {
  x <- as.numeric(x)
  y <- as.numeric(y)
  e <- resid(lm(y ~ x))
  a <- mean(y) - b_blom * mean(x)
  a + b_blom * x + e
}

moment.nohet.safe <- function(x1, x2) {
  z1 <- var(x2) - var(x1)
  z2 <- cov(x2 - x1, x1) + var(x1)
  w <- z1 / z2

  if (!is.finite(w) || abs(z2) < 1e-8) return(NA_real_)

  disc <- (2 - w)^2 + 4 * w
  if (disc < 0) return(NA_real_)

  root_plus <- (-(2 - w) + sqrt(disc)) / 2
  root_minus <- (-(2 - w) - sqrt(disc)) / 2

  b <- ifelse(cov(x1, x2) > 0, root_plus, root_minus)
  if (!is.finite(b) || abs(b) > 10) b <- NA_real_
  b
}

moment.nohet_pseudo_exact <- function(x, y, b_mom) {
  if (!is.finite(b_mom)) return(rep(NA_real_, length(x)))
  x <- as.numeric(x)
  y <- as.numeric(y)
  e <- resid(lm(y ~ x))
  a <- mean(y) - b_mom * mean(x)
  a + b_mom * x + e
}

theme_fig3_panel <- function(base_size = 8) {
  theme_minimal(base_size = base_size) +
    theme(
      plot.margin = margin(1, 1, 1, 1),
      axis.title.y = element_text(margin = margin(r = 0)),
      panel.grid.minor = element_blank()
    )
}

make_column_header <- function(labels, rel_widths, base_size = 10,
                               face = "bold", vjust = 0.5) {
  heads <- lapply(labels, function(lab) {
    cowplot::ggdraw() +
      cowplot::draw_label(
        lab,
        x = 0.5, y = 0.5,
        hjust = 0.5, vjust = vjust,
        fontface = face, size = base_size
      )
  })
  cowplot::plot_grid(plotlist = heads, ncol = length(labels), rel_widths = rel_widths)
}

text_panel_math <- function(line1, line2 = NULL, line3 = NULL, line4 = NULL,
                            line5 = NULL, line6 = NULL, line7 = NULL,
                            base_size = 8, hjust = 0.5,
                            line_y_top = 0.95, line_y_bot = 0) {
  lines <- c(line1, line2, line3, line4, line5, line6, line7)
  lines <- lines[!vapply(lines, function(z) is.null(z) || identical(z, ""), logical(1))]
  ys <- if (length(lines) == 1) 0.5 else seq(line_y_top, line_y_bot, length.out = length(lines))

  ggplot(data.frame(x = 0.5, y = ys, lab = lines), aes(x, y, label = lab)) +
    geom_text(parse = TRUE, hjust = hjust, vjust = 0.5,
              size = base_size * 0.35, lineheight = 2) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), expand = FALSE, clip = "off") +
    theme_void(base_size = base_size) +
    theme(plot.margin = margin(3, 3, 3, 3))
}

chiolero_plot_simple <- function(df, init_var, final_var, y_limits = NULL,
                                 y_label = "value", theta = NULL,
                                 var_p = NULL, seed = 123,
                                 jitter_width = 0.08) {
  init_sym <- rlang::ensym(init_var)
  final_sym <- rlang::ensym(final_var)

  set.seed(seed)
  jitter <- runif(nrow(df), -jitter_width, jitter_width)

  d <- df %>%
    mutate(
      x1 = 1 + jitter,
      x2 = 2 + jitter,
      y1 = !!init_sym,
      y2 = !!final_sym
    )

  p <- ggplot(d, aes(x = x1, y = y1)) +
    geom_segment(aes(xend = x2, yend = y2), alpha = 0.12, linewidth = 0.2) +
    geom_point(size = 1.2, alpha = 0.25) +
    geom_point(aes(x = x2, y = y2), size = 1.2, alpha = 0.25) +
    scale_x_continuous(breaks = c(1, 2), labels = c("initial", "final"),
                       expand = c(0.1, 0.1)) +
    coord_cartesian(ylim = y_limits) +
    labs(x = NULL, y = y_label) +
    theme_fig3_panel()

  if (!is.null(var_p)) {
    p <- p +
      annotate("text", x = 1.5, y = Inf, label = format_p_label(var_p),
               hjust = 0.5, vjust = 1, size = 2)
  }

  if (!is.null(theta)) {
    p <- p +
      annotate("text", x = 1.5, y = Inf,
               label = paste0("s[2]/s[1]==", round(theta, 2)),
               parse = TRUE, hjust = 0.5, vjust = 2.4, size = 2)
  }

  p
}

estimate_ci <- function(df, y_var, x_var = log_init_obs, n_boot = 10000, seed = 1) {
  y_sym <- rlang::ensym(y_var)
  x_sym <- rlang::ensym(x_var)
  fit <- lm(rlang::new_formula(rlang::expr(!!y_sym), rlang::expr(!!x_sym)), data = df)
  estimate <- unname(coef(fit)[2])

  set.seed(seed)
  boot_out <- boot::boot(
    data = df,
    statistic = function(d, i) {
      di <- d[i, , drop = FALSE]
      coef(lm(rlang::new_formula(rlang::expr(!!y_sym), rlang::expr(!!x_sym)), data = di))[2]
    },
    R = n_boot
  )

  ci <- quantile(boot_out$t, c(0.025, 0.975), na.rm = TRUE)
  c(estimate = estimate, ci_lower = ci[[1]], ci_upper = ci[[2]])
}

make_relationship_plot <- function(df, y, xlim = NULL, ylim = NULL,
                                   ylab = "change", cor_label_x = NULL,
                                   show_cor = TRUE) {
  y_sym <- rlang::ensym(y)

  p <- ggplot(df, aes(x = log_init_obs, y = !!y_sym)) +
    geom_point(alpha = 0.2, size = 0.8) +
    geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.25) +
    geom_smooth(method = "lm", se = FALSE, linewidth = 0.5) +
    coord_cartesian(xlim = xlim, ylim = ylim) +
    labs(x = NULL, y = ylab) +
    theme_fig3_panel() +
    theme(aspect.ratio = 1)

  if (show_cor) {
    p <- p +
      ggpubr::stat_cor(
        aes(label = after_stat(p.label)),
        size = 2,
        label.x = cor_label_x,
        label.y = Inf,
        hjust = 0.5,
        vjust = 1
      )
  }

  p
}

make_slope_plot_original <- function(slopes, ylim = NULL) {
  ggplot(slopes, aes(x = method, y = estimate)) +
    geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.25) +
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.05, linewidth = 0.3) +
    geom_point(size = 1.2) +
    coord_cartesian(ylim = ylim) +
    labs(x = NULL, y = "slope") +
    theme_fig3_panel() +
    theme(
      aspect.ratio = 1,
      axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
    )
}

make_slope_plot_overlay <- function(slopes, ylim = NULL) {
  ggplot(slopes, aes(x = method, y = estimate)) +
    geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.25) +
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper, color = method), width = 0.05, linewidth = 0.3) +
    geom_point(aes(color = method), size = 1.2) +
    scale_color_manual(values = method_colors, guide = "none") +
    coord_cartesian(ylim = ylim) +
    labs(x = NULL, y = "slope") +
    theme_fig3_panel() +
    theme(
      aspect.ratio = 1,
      axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
    )
}

make_overlay_plot <- function(result, xlim = NULL, ylim = NULL) {
  df <- result$data
  line_df <- tibble(
    method = factor(c("Observed", "Kelly-Price", "Blomqvist", "Moments"),
                    levels = c("Observed", "Kelly-Price", "Blomqvist", "Moments")),
    y = c("log_change_obs", "log_change_kp_adj", "log_change_blomq_adj", "log_change_mom_adj")
  ) %>%
    mutate(
      fit = map(y, ~ lm(reformulate("log_init_obs", .x), data = df)),
      intercept = map_dbl(fit, ~ coef(.x)[1]),
      slope = map_dbl(fit, ~ coef(.x)[2])
    )

  ggplot(df, aes(x = log_init_obs, y = log_change_obs)) +
    geom_point(alpha = 0.06, size = 0.55) +
    geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.25) +
    geom_abline(
      data = line_df,
      aes(intercept = intercept, slope = slope, color = method, linetype = method),
      linewidth = 0.6
    ) +
    coord_cartesian(xlim = xlim, ylim = ylim) +
    scale_color_manual(values = method_colors) +
    scale_linetype_manual(values = method_linetypes) +
    labs(x = NULL, y = "observed change") +
    theme_fig3_panel() +
    theme(legend.position = "none", aspect.ratio = 1)
}
```

# Prepare Case Studies

```{r prepare-case-studies}
prepare_pdam <- function(n_boot = 10000) {
  pdam0 <- read_csv("data/PdamRbleaching.csv", show_col_types = FALSE)
  pdam <- pdam0 %>%
    mutate(
      log_init_raw = log(juntotal),
      log_final_raw = log(augtotal),
      log_change_raw = log_final_raw - log_init_raw
    )

  grand_mean_init <- mean(pdam$log_init_raw)
  grand_mean_final <- mean(pdam$log_final_raw)

  pdam <- pdam %>%
    group_by(sym) %>%
    mutate(
      group_resid_init = log_init_raw - mean(log_init_raw),
      group_resid_final = log_final_raw - mean(log_final_raw)
    ) %>%
    ungroup() %>%
    mutate(
      log_init_obs = grand_mean_init + group_resid_init,
      log_final_obs = grand_mean_final + group_resid_final,
      log_change_obs = log_final_obs - log_init_obs
    ) %>%
    select(colony, sym, log_init_obs, log_final_obs, log_change_obs)

  var_x_abs <- var(pdam$log_init_obs, na.rm = TRUE)
  k <- 0.4648801
  v_err_abs <- k * var_x_abs
  v_pop <- (1 - k) * var_x_abs

  kp_full <- rttm.adj(pdam$log_init_obs, pdam$log_final_obs)
  obs_ci <- estimate_ci(pdam, log_change_obs, n_boot = n_boot, seed = 1)
  kp_ci <- {
    pdam_kp <- pdam %>% mutate(log_change_kp_adj = kp_full$adj.growth)
    estimate_ci(pdam_kp, log_change_kp_adj, n_boot = n_boot, seed = 2)
  }
  b_blomq <- blomqvist_b(obs_ci[["estimate"]], k)
  blomq_ci <- c(
    estimate = b_blomq,
    ci_lower = blomqvist_b(obs_ci[["ci_lower"]], k),
    ci_upper = blomqvist_b(obs_ci[["ci_upper"]], k)
  )
  b_mom <- moment.nohet.safe(pdam$log_init_obs, pdam$log_final_obs)

  set.seed(3)
  mom_boot <- boot::boot(
    data = pdam,
    statistic = function(d, i) {
      di <- d[i, , drop = FALSE]
      moment.nohet.safe(di$log_init_obs, di$log_final_obs)
    },
    R = n_boot
  )
  mom_ci <- quantile(mom_boot$t, c(0.025, 0.975), na.rm = TRUE)

  pdam <- pdam %>%
    mutate(
      log_change_kp_adj = kp_full$adj.growth,
      log_change_blomq_adj = blomqvist_pseudo_exact(log_init_obs, log_change_obs, b_blomq),
      log_change_mom_adj = moment.nohet_pseudo_exact(log_init_obs, log_change_obs, b_mom)
    )

  v_ind <- var(pdam$log_change_obs) - (2 * v_err_abs) - (b_blomq^2 * v_pop)

  slopes <- tibble(
    method = factor(c("Obs.", "K.-P.", "Blom.", "Mom."),
                    levels = c("Obs.", "K.-P.", "Blom.", "Mom.")),
    estimate = c(obs_ci[["estimate"]], kp_ci[["estimate"]], b_blomq, b_mom),
    ci_lower = c(obs_ci[["ci_lower"]], kp_ci[["ci_lower"]], blomq_ci[["ci_lower"]], mom_ci[[1]]),
    ci_upper = c(obs_ci[["ci_upper"]], kp_ci[["ci_upper"]], blomq_ci[["ci_upper"]], mom_ci[[2]])
  )

  list(
    data = pdam,
    slopes = slopes,
    trait_ylim = c(-4.1, -0.5),
    change_ylim = c(range(pdam$log_change_obs, na.rm = TRUE)[1],
                    range(pdam$log_change_obs, na.rm = TRUE)[2] + 0.1),
    xlim = c(-3.2, -1.0),
    slope_ylim = c(-1.2, 0.2),
    theta = sd(pdam$log_final_obs) / sd(pdam$log_init_obs),
    var_p = var.test(pdam$log_init_obs, pdam$log_final_obs, paired = TRUE)$p.value,
    cor_label_x = -2.25,
    y_label = "log S/H",
    text = list(
      line1 = "bold('Bleaching in')",
      line2 = "bolditalic(Pocillopora)",
      line3 = "bolditalic(damicornis)",
      line4 = sprintf("n==bold(%s)", formatC(nrow(pdam), format = "fg")),
      line5 = sprintf("s[X]^2==bold(%s)", formatC(v_pop, format = "fg", digits = 3)),
      line6 = sprintf("s[Delta]^2==bold(%s)", formatC(v_ind, format = "fg", digits = 3)),
      line7 = sprintf("s[err]^2==bold(%s)", formatC(v_err_abs, format = "fg", digits = 3))
    )
  )
}

prepare_mill <- function(n_boot = 10000) {
  mill <- read_csv("data/subMil1.csv", show_col_types = FALSE) %>%
    filter(size > 0, InitialSize > 1) %>%
    mutate(
      log_init_obs = log(InitialSize),
      log_final_obs = log(size),
      log_change_obs = log_final_obs - log_init_obs
    )

  var_x_abs <- var(mill$log_init_obs, na.rm = TRUE)
  v_err_abs <- 0.01301945
  v_pop <- var_x_abs - v_err_abs
  k <- v_err_abs / var_x_abs

  kp_full <- rttm.adj(mill$log_init_obs, mill$log_final_obs)
  obs_ci <- estimate_ci(mill, log_change_obs, n_boot = n_boot, seed = 11)
  kp_ci <- {
    mill_kp <- mill %>% mutate(log_change_kp_adj = kp_full$adj.growth)
    estimate_ci(mill_kp, log_change_kp_adj, n_boot = n_boot, seed = 12)
  }
  b_blomq <- blomqvist_b(obs_ci[["estimate"]], k)
  blomq_ci <- c(
    estimate = b_blomq,
    ci_lower = blomqvist_b(obs_ci[["ci_lower"]], k),
    ci_upper = blomqvist_b(obs_ci[["ci_upper"]], k)
  )
  b_mom <- moment.nohet.safe(mill$log_init_obs, mill$log_final_obs)

  set.seed(13)
  mom_boot <- boot::boot(
    data = mill,
    statistic = function(d, i) {
      di <- d[i, , drop = FALSE]
      moment.nohet.safe(di$log_init_obs, di$log_final_obs)
    },
    R = n_boot
  )
  mom_ci <- quantile(mom_boot$t, c(0.025, 0.975), na.rm = TRUE)

  mill <- mill %>%
    mutate(
      log_change_kp_adj = kp_full$adj.growth,
      log_change_blomq_adj = blomqvist_pseudo_exact(log_init_obs, log_change_obs, b_blomq),
      log_change_mom_adj = moment.nohet_pseudo_exact(log_init_obs, log_change_obs, b_mom)
    )

  v_ind <- var(mill$log_change_obs) - (2 * v_err_abs) - (b_blomq^2 * v_pop)

  slopes <- tibble(
    method = factor(c("Obs.", "K.-P.", "Blom.", "Mom."),
                    levels = c("Obs.", "K.-P.", "Blom.", "Mom.")),
    estimate = c(obs_ci[["estimate"]], kp_ci[["estimate"]], b_blomq, b_mom),
    ci_lower = c(obs_ci[["ci_lower"]], kp_ci[["ci_lower"]], blomq_ci[["ci_lower"]], mom_ci[[1]]),
    ci_upper = c(obs_ci[["ci_upper"]], kp_ci[["ci_upper"]], blomq_ci[["ci_upper"]], mom_ci[[2]])
  )

  trait_range <- range(c(mill$log_init_obs, mill$log_final_obs), na.rm = TRUE)
  change_range <- range(mill$log_change_obs, na.rm = TRUE)

  list(
    data = mill,
    slopes = slopes,
    trait_ylim = c(-0.5, 6),
    change_ylim = change_range,
    xlim = trait_range,
    slope_ylim = NULL,
    theta = sd(mill$log_final_obs) / sd(mill$log_init_obs),
    var_p = var.test(mill$log_init_obs, mill$log_final_obs, paired = TRUE)$p.value,
    cor_label_x = 2.25,
    y_label = "log size (cm)",
    text = list(
      line1 = "bold('Growth in')",
      line2 = "bolditalic(Acropora)",
      line3 = "bolditalic(cervicornis)",
      line4 = sprintf("n==bold(%s)", formatC(nrow(mill), format = "fg")),
      line5 = sprintf("s[X]^2==bold(%s)", formatC(v_pop, format = "fg", digits = 3)),
      line6 = sprintf("s[Delta]^2==bold(%s)", formatC(v_ind, format = "fg", digits = 3)),
      line7 = sprintf("s[err]^2==bold(%s)", formatC(v_err_abs, format = "fg", digits = 3))
    )
  )
}

pdam_result <- prepare_pdam()
mill_result <- prepare_mill()
case_results <- list(pdam_result, mill_result)
```

# Row Constructors

```{r row-constructors}
case_text_panel <- function(result) {
  do.call(text_panel_math, result$text)
}

make_original_row <- function(result) {
  df <- result$data
  cowplot::plot_grid(
    case_text_panel(result),
    chiolero_plot_simple(df, log_init_obs, log_final_obs,
                         y_limits = result$trait_ylim, y_label = result$y_label,
                         theta = result$theta, var_p = result$var_p),
    make_relationship_plot(df, log_change_obs, xlim = result$xlim,
                           ylim = result$change_ylim, cor_label_x = result$cor_label_x),
    make_relationship_plot(df, log_change_kp_adj, xlim = result$xlim,
                           ylim = result$change_ylim, cor_label_x = result$cor_label_x),
    make_relationship_plot(df, log_change_blomq_adj, xlim = result$xlim,
                           ylim = result$change_ylim, cor_label_x = result$cor_label_x),
    make_relationship_plot(df, log_change_mom_adj, xlim = result$xlim,
                           ylim = result$change_ylim, cor_label_x = result$cor_label_x),
    make_slope_plot_original(result$slopes, ylim = result$slope_ylim),
    nrow = 1,
    align = "h",
    rel_widths = c(1, 0.6, 1, 1, 1, 1, 1)
  )
}

make_drop_adjusted_row <- function(result) {
  df <- result$data
  cowplot::plot_grid(
    case_text_panel(result),
    chiolero_plot_simple(df, log_init_obs, log_final_obs,
                         y_limits = result$trait_ylim, y_label = result$y_label,
                         theta = result$theta, var_p = result$var_p),
    make_relationship_plot(df, log_change_obs, xlim = result$xlim,
                           ylim = result$change_ylim, cor_label_x = result$cor_label_x),
    make_slope_plot_original(result$slopes, ylim = result$slope_ylim),
    nrow = 1,
    align = "h",
    rel_widths = c(1, 0.6, 0.8, 0.8)
  )
}

make_minimal_row <- function(result) {
  df <- result$data
  cowplot::plot_grid(
    case_text_panel(result),
    chiolero_plot_simple(df, log_init_obs, log_final_obs,
                         y_limits = result$trait_ylim, y_label = result$y_label,
                         theta = result$theta, var_p = result$var_p),
    make_slope_plot_original(result$slopes, ylim = result$slope_ylim),
    nrow = 1,
    align = "h",
    rel_widths = c(1, 0.6, 0.8)
  )
}

make_overlay_row <- function(result) {
  df <- result$data
  cowplot::plot_grid(
    case_text_panel(result),
    chiolero_plot_simple(df, log_init_obs, log_final_obs,
                         y_limits = result$trait_ylim, y_label = result$y_label,
                         theta = result$theta, var_p = result$var_p),
    make_overlay_plot(result, xlim = result$xlim, ylim = result$change_ylim),
    make_slope_plot_overlay(result$slopes, ylim = result$slope_ylim),
    nrow = 1,
    align = "h",
    rel_widths = c(1, 0.6, 0.8, 0.8)
  )
}
```

# Original Recreated

```{r fig3-original, fig.width=8, fig.height=3.2}
original_header <- make_column_header(
  labels = c("1. Dataset", "2. Observed\nvalues", "3. Observed\nbaseline effect",
             "4. Kelly-Price\nadjusted", "5. Blomqvist\nadjusted",
             "6. Moments\nadjusted", "7. Slope\ncomparison"),
  rel_widths = c(1, 0.6, 1, 1, 1, 1, 1),
  base_size = 10
)

original_footer <- make_column_header(
  labels = c("", "timepoint", "initial", "initial", "initial", "initial", "method"),
  rel_widths = c(1, 0.6, 1, 1, 1, 1, 1),
  base_size = 8,
  face = "plain",
  vjust = -0.2
)

fig3_original_recreated <- cowplot::plot_grid(
  plotlist = c(list(original_header), map(case_results, make_original_row), list(original_footer)),
  ncol = 1,
  labels = c("", "A", "B", ""),
  rel_heights = c(0.25, 1, 1, 0.08)
)

fig3_original_recreated

ggsave(
  filename = "figures/figure3_alternatives/fig3_original_recreated.png",
  plot = fig3_original_recreated,
  width = 200,
  height = 80,
  units = "mm",
  bg = "white"
)
```

# Alternative 1: Original Without Adjusted Relationship Panels

```{r fig3-drop-adjusted, fig.width=5.5, fig.height=3.2}
drop_header <- make_column_header(
  labels = c("1. Dataset", "2. Observed\nvalues",
             "3. Observed\nbaseline effect", "4. Slope\ncomparison"),
  rel_widths = c(1, 0.6, 0.8, 0.8),
  base_size = 10
)

drop_footer <- make_column_header(
  labels = c("", "timepoint", "initial", "method"),
  rel_widths = c(1, 0.6, 0.8, 0.8),
  base_size = 8,
  face = "plain",
  vjust = -0.2
)

fig3_alt_drop_adjusted_panels <- cowplot::plot_grid(
  plotlist = c(list(drop_header), map(case_results, make_drop_adjusted_row), list(drop_footer)),
  ncol = 1,
  labels = c("", "A", "B", ""),
  rel_heights = c(0.25, 1, 1, 0.08)
)

fig3_alt_drop_adjusted_panels

ggsave(
  filename = "figures/figure3_alternatives/fig3_alt_drop_adjusted_panels.png",
  plot = fig3_alt_drop_adjusted_panels,
  width = 145,
  height = 80,
  units = "mm",
  bg = "white"
)
```

# Alternative 2: Reviewer-Minimal

```{r fig3-reviewer-minimal, fig.width=4.5, fig.height=3.2}
minimal_header <- make_column_header(
  labels = c("1. Dataset", "2. Observed\nvalues", "3. Slope\ncomparison"),
  rel_widths = c(1, 0.6, 0.8),
  base_size = 10
)

minimal_footer <- make_column_header(
  labels = c("", "timepoint", "method"),
  rel_widths = c(1, 0.6, 0.8),
  base_size = 8,
  face = "plain",
  vjust = -0.2
)

fig3_alt_reviewer_minimal <- cowplot::plot_grid(
  plotlist = c(list(minimal_header), map(case_results, make_minimal_row), list(minimal_footer)),
  ncol = 1,
  labels = c("", "A", "B", ""),
  rel_heights = c(0.25, 1, 1, 0.08)
)

fig3_alt_reviewer_minimal

ggsave(
  filename = "figures/figure3_alternatives/fig3_alt_reviewer_minimal.png",
  plot = fig3_alt_reviewer_minimal,
  width = 115,
  height = 80,
  units = "mm",
  bg = "white"
)
```

# Alternative 3: Compact Overlay

```{r fig3-overlay, fig.width=5.5, fig.height=3.2}
overlay_header <- make_column_header(
  labels = c("1. Dataset", "2. Observed\nvalues",
             "3. Baseline-change", "4. Slopes"),
  rel_widths = c(1, 0.6, 0.8, 0.8),
  base_size = 10
)

overlay_footer <- make_column_header(
  labels = c("", "timepoint", "initial", "method"),
  rel_widths = c(1, 0.6, 0.8, 0.8),
  base_size = 8,
  face = "plain",
  vjust = -0.2
)

fig3_alt_overlay <- cowplot::plot_grid(
  plotlist = c(list(overlay_header), map(case_results, make_overlay_row), list(overlay_footer)),
  ncol = 1,
  labels = c("", "A", "B", ""),
  rel_heights = c(0.25, 1, 1, 0.08)
)

fig3_alt_overlay

ggsave(
  filename = "figures/figure3_alternatives/fig3_alt_overlay.png",
  plot = fig3_alt_overlay,
  width = 145,
  height = 80,
  units = "mm",
  bg = "white"
)
```
