---
title: "Figure 1 alternatives for reviewer response"
date: "May 2026"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE)
```

# Purpose

Reviewer feedback suggested moving the separate baseline-change correction panels
from Figure 1 columns 4-7 to the supplement because the slope-estimate panel
conveys the same inferential result with less clutter.

This notebook recreates the submitted Figure 1 and constructs several alternatives
that keep different amounts of the baseline-change visualization.

The preferred organization is to keep this file as a figure-design sandbox. The
simulation data are regenerated here from the same parameter values and seeds,
rather than saved out of `RTM_analysis.Rmd`, so this file can be rendered
independently while leaving the main manuscript analysis untouched.

# Setup

```{r libraries}
library(tidyverse)
library(cowplot)
library(ggpubr)
library(boot)
```

```{r helpers}
dir.create("figures/figure1_alternatives", recursive = TRUE, showWarnings = FALSE)

method_colors <- c(
  "Observed" = "#0072B2",
  "Kelly-Price" = "#D55E00",
  "Blomqvist" = "#009E73",
  "Moments" = "#CC79A7",
  "Obs." = "#0072B2",
  "K.-P." = "#D55E00",
  "Blom." = "#009E73",
  "Mom." = "#CC79A7"
)

true_slope_color <- "#8B0000"

method_linetypes <- c(
  "Observed" = "solid",
  "Kelly-Price" = "22",
  "Blomqvist" = "42",
  "Moments" = "73"
)

rttm.adj <- function(m1, m2) {
  raw.growth <- m2 - m1
  vart <- var.test(m1, m2, paired = TRUE)
  vpv <- vart$p.value
  rho <- cor.test(m1, m2)$estimate
  m1sd <- sd(m1)
  m2sd <- sd(m2)
  m1v <- var(m1)
  m2v <- var(m2)
  m1m <- mean(m1)
  m2m <- mean(m2)
  pm <- mean(raw.growth)
  rho2 <- (2 * rho * m1sd * m2sd) / (m1v + m2v)
  rhof <- ifelse(vpv <= 0.05, rho, rho2)
  dstar <- (rhof * (m1 - m1m) - (m2 - m2m)) * -1
  data.frame(raw.growth = raw.growth, dstar = dstar, adj.growth = pm + dstar)
}

blomqvist_b <- function(b_obs, k) {
  (b_obs + k) / (1 - k)
}

blomqvist_pseudo_exact <- function(x, y, b_blom) {
  x <- as.numeric(x)
  y <- as.numeric(y)
  e <- resid(lm(y ~ x))
  a <- mean(y) - b_blom * mean(x)
  a + b_blom * x + e
}

moment.nohet.safe <- function(x1, x2) {
  z1 <- var(x2) - var(x1)
  z2 <- cov(x2 - x1, x1) + var(x1)
  w <- z1 / z2

  if (!is.finite(w) || abs(z2) < 1e-8) return(NA_real_)

  disc <- (2 - w)^2 + 4 * w
  if (disc < 0) return(NA_real_)

  root_plus <- (-(2 - w) + sqrt(disc)) / 2
  root_minus <- (-(2 - w) - sqrt(disc)) / 2

  b <- ifelse(cov(x1, x2) > 0, root_plus, root_minus)
  if (!is.finite(b) || abs(b) > 10) b <- NA_real_
  b
}

moment.nohet_pseudo_exact <- function(x, y, b_mom) {
  if (!is.finite(b_mom)) return(rep(NA_real_, length(x)))
  x <- as.numeric(x)
  y <- as.numeric(y)
  e <- resid(lm(y ~ x))
  a <- mean(y) - b_mom * mean(x)
  a + b_mom * x + e
}

theme_fig1_panel <- function(base_size = 8) {
  theme_minimal(base_size = base_size) +
    theme(
      plot.margin = margin(1, 1, 1, 1),
      axis.title.y = element_text(margin = margin(r = -2)),
      panel.grid.minor = element_blank()
    )
}

make_column_header <- function(labels, rel_widths, base_size = 10,
                               face = "bold", vjust = 0.5) {
  heads <- lapply(labels, function(lab) {
    cowplot::ggdraw() +
      cowplot::draw_label(
        lab,
        x = 0.5, y = 0.5,
        hjust = 0.5, vjust = vjust,
        fontface = face, size = base_size
      )
  })
  cowplot::plot_grid(plotlist = heads, ncol = length(labels), rel_widths = rel_widths)
}

text_panel_math <- function(line1, line2 = NULL, line3 = NULL,
                            base_size = 8, line_y_top = 0.78,
                            line_y_bot = 0.22) {
  lines <- c(line1, line2, line3)
  lines <- lines[!vapply(lines, function(z) is.null(z) || identical(z, ""), logical(1))]
  ys <- if (length(lines) == 1) 0.5 else seq(line_y_top, line_y_bot, length.out = length(lines))

  ggplot(data.frame(x = 0.5, y = ys, lab = lines), aes(x, y, label = lab)) +
    geom_text(parse = TRUE, hjust = 0.5, vjust = 0.5,
              size = base_size * 0.35, fontface = "bold") +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), expand = FALSE, clip = "off") +
    theme_void(base_size = base_size) +
    theme(plot.margin = margin(3, 3, 3, 3))
}

chiolero_plot_simple <- function(df, init_var, final_var, seed = 123,
                                 jitter_width = 0.08) {
  init_sym <- rlang::ensym(init_var)
  final_sym <- rlang::ensym(final_var)

  set.seed(seed)
  jitter <- runif(nrow(df), -jitter_width, jitter_width)

  d <- df %>%
    mutate(
      x1 = 1 + jitter,
      x2 = 2 + jitter,
      y1 = !!init_sym,
      y2 = !!final_sym
    )

  ggplot(d, aes(x = x1, y = y1)) +
    geom_segment(aes(xend = x2, yend = y2), alpha = 0.12, linewidth = 0.2) +
    geom_point(size = 1.2, alpha = 0.25) +
    geom_point(aes(x = x2, y = y2), size = 1.2, alpha = 0.25) +
    scale_x_continuous(breaks = c(1, 2), labels = c("initial", "final"),
                       expand = c(0.1, 0.1)) +
    labs(x = NULL, y = "value") +
    theme_fig1_panel()
}

estimate_ci <- function(df, y_var, x_var = log_init_obs, n_boot = 10000, seed = 1) {
  y_sym <- rlang::ensym(y_var)
  x_sym <- rlang::ensym(x_var)
  fit <- lm(rlang::new_formula(rlang::expr(!!y_sym), rlang::expr(!!x_sym)), data = df)
  estimate <- coef(fit)[2]

  set.seed(seed)
  boot_out <- boot::boot(
    data = df,
    statistic = function(d, i) {
      di <- d[i, , drop = FALSE]
      coef(lm(rlang::new_formula(rlang::expr(!!y_sym), rlang::expr(!!x_sym)), data = di))[2]
    },
    R = n_boot
  )

  ci <- quantile(boot_out$t, c(0.025, 0.975), na.rm = TRUE)
  c(estimate = unname(estimate), ci_lower = ci[[1]], ci_upper = ci[[2]])
}

run_simulation_scenario <- function(n, b, v_ind, v_err, v_pop = 1,
                                    seed = 4, n_boot = 10000) {
  set.seed(seed)

  df <- tibble(
    log_init_true = rnorm(n, 0, sqrt(v_pop)),
    log_change_true = b * log_init_true + rnorm(n, 0, sqrt(v_ind)),
    log_final_true = log_init_true + log_change_true,
    e_init = rnorm(n, 0, sqrt(v_err)),
    e_final = rnorm(n, 0, sqrt(v_err)),
    log_init_obs = log_init_true + e_init,
    log_final_obs = log_final_true + e_final,
    log_change_obs = log_final_obs - log_init_obs
  )

  kp <- rttm.adj(df$log_init_obs, df$log_final_obs)
  b_obs <- coef(lm(log_change_obs ~ log_init_obs, data = df))[2]
  k <- v_err / (v_pop + v_err)
  b_blomq <- blomqvist_b(b_obs, k)
  b_mom <- moment.nohet.safe(df$log_init_obs, df$log_final_obs)

  df <- df %>%
    mutate(
      log_change_kp_adj = kp$adj.growth,
      log_change_blomq_adj = blomqvist_pseudo_exact(log_init_obs, log_change_obs, b_blomq),
      log_change_mom_adj = moment.nohet_pseudo_exact(log_init_obs, log_change_obs, b_mom)
    )

  obs_ci <- estimate_ci(df, log_change_obs, n_boot = n_boot, seed = seed + 10)
  kp_ci <- estimate_ci(df, log_change_kp_adj, n_boot = n_boot, seed = seed + 20)
  blomq_ci <- c(
    estimate = b_blomq,
    ci_lower = blomqvist_b(obs_ci[["ci_lower"]], k),
    ci_upper = blomqvist_b(obs_ci[["ci_upper"]], k)
  )

  set.seed(seed + 30)
  mom_boot <- boot::boot(
    data = df,
    statistic = function(d, i) {
      di <- d[i, , drop = FALSE]
      moment.nohet.safe(di$log_init_obs, di$log_final_obs)
    },
    R = n_boot
  )
  mom_ci <- quantile(mom_boot$t, c(0.025, 0.975), na.rm = TRUE)

  slopes <- tibble(
    method = factor(c("Obs.", "K.-P.", "Blom.", "Mom."),
                    levels = c("Obs.", "K.-P.", "Blom.", "Mom.")),
    estimate = c(obs_ci[["estimate"]], kp_ci[["estimate"]], b_blomq, b_mom),
    ci_lower = c(obs_ci[["ci_lower"]], kp_ci[["ci_lower"]], blomq_ci[["ci_lower"]], mom_ci[[1]]),
    ci_upper = c(obs_ci[["ci_upper"]], kp_ci[["ci_upper"]], blomq_ci[["ci_upper"]], mom_ci[[2]])
  )

  list(
    data = df,
    slopes = slopes,
    params = list(b = b, v_ind = v_ind, v_err = v_err, v_pop = v_pop)
  )
}

make_relationship_plot <- function(df, y, title = NULL, xlim = NULL,
                                   ylim = NULL, ylab = "change") {
  y_sym <- rlang::ensym(y)

  ggplot(df, aes(x = log_init_obs, y = !!y_sym)) +
    geom_point(alpha = 0.2, size = 0.7) +
    geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.25) +
    geom_smooth(method = "lm", se = FALSE, linewidth = 0.5) +
    coord_cartesian(xlim = xlim, ylim = ylim) +
    labs(title = title, x = NULL, y = ylab) +
    theme_fig1_panel() +
    theme(aspect.ratio = 1)
}

make_slope_plot <- function(slopes, b_true, ylim = c(-0.61, 0.61)) {
  ggplot(slopes, aes(x = method, y = estimate)) +
    geom_hline(yintercept = b_true, color = true_slope_color, linewidth = 0.3) +
    geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.25) +
    geom_errorbar(
      aes(ymin = ci_lower, ymax = ci_upper, color = method),
      width = 0.05,
      linewidth = 0.3
    ) +
    geom_point(aes(color = method), size = 1.2) +
    scale_color_manual(values = method_colors, guide = "none") +
    coord_cartesian(ylim = ylim, clip = "off") +
    labs(x = NULL, y = "slope") +
    theme_fig1_panel() +
    theme(
      aspect.ratio = 1,
      axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
    )
}

make_slope_plot_original <- function(slopes, b_true, ylim = c(-0.61, 0.61)) {
  ggplot(slopes, aes(x = method, y = estimate)) +
    geom_hline(yintercept = b_true, color = "red", linewidth = 0.3) +
    geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.25) +
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.05, linewidth = 0.3) +
    geom_point(size = 1.2) +
    coord_cartesian(ylim = ylim, clip = "off") +
    labs(x = NULL, y = "slope") +
    theme_fig1_panel() +
    theme(
      aspect.ratio = 1,
      axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
    )
}

make_overlay_plot <- function(result, xlim = NULL, ylim = NULL) {
  df <- result$data
  line_df <- tibble(
    method = factor(c("Observed", "Kelly-Price", "Blomqvist", "Moments"),
                    levels = c("Observed", "Kelly-Price", "Blomqvist", "Moments")),
    y = c("log_change_obs", "log_change_kp_adj", "log_change_blomq_adj", "log_change_mom_adj")
  ) %>%
    mutate(
      fit = map(y, ~ lm(reformulate("log_init_obs", .x), data = df)),
      intercept = map_dbl(fit, ~ coef(.x)[1]),
      slope = map_dbl(fit, ~ coef(.x)[2])
    )

  ggplot(df, aes(x = log_init_obs, y = log_change_obs)) +
    geom_point(alpha = 0.06, size = 0.55) +
    geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.25) +
    geom_abline(
      data = line_df,
      aes(intercept = intercept, slope = slope, color = method, linetype = method),
      linewidth = 0.6
    ) +
    geom_abline(intercept = 0, slope = result$params$b, color = true_slope_color, linewidth = 0.45) +
    coord_cartesian(xlim = xlim, ylim = ylim) +
    scale_color_manual(values = method_colors) +
    scale_linetype_manual(values = method_linetypes) +
    labs(x = NULL, y = "observed change", color = "Method", linetype = "Method") +
    theme_fig1_panel() +
    theme(legend.position = "none", aspect.ratio = 1)
}

make_method_legend <- function(base_size = 8) {
  legend_df <- tibble(
    method = factor(c("Observed", "Kelly-Price", "Blomqvist", "Moments"),
                    levels = c("Observed", "Kelly-Price", "Blomqvist", "Moments")),
    x = c(0.18, 0.57, 0.18, 0.57),
    xend = x + 0.08,
    y = c(0.7, 0.7, 0.25, 0.25)
  )

  ggplot(legend_df) +
    geom_segment(
      aes(x = x, xend = xend, y = y, yend = y, color = method, linetype = method),
      linewidth = 0.65
    ) +
    geom_text(aes(x = xend + 0.015, y = y, label = method),
              hjust = 0, vjust = 0.5, size = base_size * 0.35) +
    scale_color_manual(values = method_colors) +
    scale_linetype_manual(values = method_linetypes) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), expand = FALSE, clip = "off") +
    theme_void(base_size = base_size) +
    theme(legend.position = "none", plot.margin = margin(0, 0, 0, 0))
}
```

# Simulate Scenarios

These are the four rows used in the submitted Figure 1.

```{r simulate}
n <- 200
v_pop <- 1

scenario_specs <- tribble(
  ~scenario, ~b,   ~v_ind, ~v_err,
  "A",       0,    0,      0.5,
  "B",       0,    1,      0,
  "C",      -0.3,  0,      0.5,
  "D",      -0.3,  1,      0
)

scenario_results <- scenario_specs %>%
  mutate(result = pmap(
    list(b, v_ind, v_err),
    ~ run_simulation_scenario(n = n, b = ..1, v_ind = ..2, v_err = ..3, v_pop = v_pop)
  ))
```

# Row Constructors

```{r row-constructors}
row_limits <- function(result) {
  df <- result$data
  max_abs_change <- df %>%
    select(log_init_true, log_change_true, log_init_obs, log_change_obs,
           log_change_kp_adj, log_change_blomq_adj, log_change_mom_adj) %>%
    summarise(max_abs = max(abs(c_across(everything())), na.rm = TRUE)) %>%
    pull(max_abs)

  max_abs_value <- df %>%
    select(log_init_true, log_final_true, log_init_obs, log_final_obs) %>%
    summarise(max_abs = max(abs(c_across(everything())), na.rm = TRUE)) %>%
    pull(max_abs)

  list(
    value = c(-max_abs_value, max_abs_value),
    change = c(-max_abs_change, max_abs_change)
  )
}

param_panel <- function(result) {
  p <- result$params
  text_panel_math(
    line1 = sprintf("b[true]==%.3g", p$b),
    line2 = sprintf("sigma[Delta]^2==%.3g", p$v_ind),
    line3 = sprintf("sigma[err]^2==%.3g", p$v_err),
    line_y_top = 0.68,
    line_y_bot = 0.2
  )
}

make_original_row <- function(result) {
  lim <- row_limits(result)
  df <- result$data

  p_true <- chiolero_plot_simple(df, log_init_true, log_final_true) +
    coord_cartesian(ylim = lim$value)
  p_obs_time <- chiolero_plot_simple(df, log_init_obs, log_final_obs) +
    coord_cartesian(ylim = lim$value)
  p_obs <- make_relationship_plot(df, log_change_obs, xlim = lim$change, ylim = lim$change)
  p_kp <- make_relationship_plot(df, log_change_kp_adj, xlim = lim$change, ylim = lim$change)
  p_blomq <- make_relationship_plot(df, log_change_blomq_adj, xlim = lim$change, ylim = lim$change)
  p_mom <- make_relationship_plot(df, log_change_mom_adj, xlim = lim$change, ylim = lim$change)
  p_slopes <- make_slope_plot_original(result$slopes, result$params$b)

  cowplot::plot_grid(
    param_panel(result), p_true, p_obs_time, p_obs, p_kp, p_blomq, p_mom, p_slopes,
    nrow = 1,
    align = "h",
    rel_widths = c(0.6, 0.6, 0.6, rep(1, 5))
  )
}

make_original_drop_adjusted_row <- function(result) {
  lim <- row_limits(result)
  df <- result$data

  p_true <- chiolero_plot_simple(df, log_init_true, log_final_true) +
    coord_cartesian(ylim = lim$value)
  p_obs_time <- chiolero_plot_simple(df, log_init_obs, log_final_obs) +
    coord_cartesian(ylim = lim$value)
  p_obs <- make_relationship_plot(df, log_change_obs, xlim = lim$change, ylim = lim$change)
  p_slopes <- make_slope_plot_original(result$slopes, result$params$b)

  cowplot::plot_grid(
    param_panel(result), p_true, p_obs_time, p_obs, p_slopes,
    nrow = 1,
    align = "h",
    rel_widths = c(0.6, 0.6, 0.6, 0.67, 0.67)
  )
}

make_reviewer_minimal_row <- function(result) {
  lim <- row_limits(result)
  df <- result$data

  cowplot::plot_grid(
    param_panel(result),
    chiolero_plot_simple(df, log_init_true, log_final_true) + coord_cartesian(ylim = lim$value),
    chiolero_plot_simple(df, log_init_obs, log_final_obs) + coord_cartesian(ylim = lim$value),
    make_slope_plot_original(result$slopes, result$params$b),
    nrow = 1,
    align = "h",
    rel_widths = c(0.5, 0.5, 0.5, 0.85)
  )
}

make_overlay_row <- function(result) {
  lim <- row_limits(result)
  df <- result$data

  cowplot::plot_grid(
    param_panel(result),
    chiolero_plot_simple(df, log_init_true, log_final_true) + coord_cartesian(ylim = lim$value),
    chiolero_plot_simple(df, log_init_obs, log_final_obs) + coord_cartesian(ylim = lim$value),
    make_overlay_plot(result, xlim = lim$change, ylim = lim$change),
    make_slope_plot(result$slopes, result$params$b),
    nrow = 1,
    align = "h",
    rel_widths = c(0.5, 0.5, 0.5, 0.78, 0.78)
  )
}
```

# Original Figure 1

```{r original-figure-1, fig.width=10, fig.height=7}
original_header <- make_column_header(
  labels = c("1. Param-\neters", "2. True\nvalues", "3. Observed\nvalues",
             "4. Observed\nbaseline effect", "5. Kelly-Price\nadjusted",
             "6. Blomqvist\nadjusted", "7. Moments\nadjusted",
             "8. Slope\ncomparison"),
  rel_widths = c(0.6, 0.6, 0.6, rep(1, 5))
)

original_footer <- make_column_header(
  labels = c("", "timepoint", "timepoint", "initial", "initial",
             "initial", "initial", "method"),
  rel_widths = c(0.6, 0.6, 0.6, rep(1, 5)),
  base_size = 8,
  face = "plain",
  vjust = -0.25
)

fig1_original_recreated <- cowplot::plot_grid(
  plotlist = c(
    list(original_header),
    scenario_results$result %>% map(make_original_row),
    list(original_footer)
  ),
  labels = c("", scenario_results$scenario, ""),
  ncol = 1,
  rel_heights = c(0.25, rep(1, nrow(scenario_results)), 0.08)
)

fig1_original_recreated

ggsave(
  filename = "figures/figure1_alternatives/fig1_original_recreated.png",
  plot = fig1_original_recreated,
  width = 200,
  height = 140,
  units = "mm",
  bg = "white"
)
```

# Alternative 1: Original Without Adjusted Relationship Panels

This version is exactly the submitted Figure 1 structure except that the three
separate adjusted baseline-change columns are removed. It keeps the observed
baseline-change relationship and the slope comparison.

```{r alternative-drop-adjusted-panels, fig.width=8, fig.height=7}
drop_adjusted_header <- make_column_header(
  labels = c("1. Param-\neters", "2. True\nvalues", "3. Observed\nvalues",
             "4. Observed\nbaseline effect", "5. Slope\ncomparison"),
  rel_widths = c(0.6, 0.6, 0.6, 0.67, 0.67),
  base_size = 10
)

drop_adjusted_footer <- make_column_header(
  labels = c("", "timepoint", "timepoint", "initial", "method"),
  rel_widths = c(0.6, 0.6, 0.6, 0.67, 0.67),
  base_size = 8,
  face = "plain",
  vjust = -0.25
)

fig1_alt_drop_adjusted_panels <- cowplot::plot_grid(
  plotlist = c(
    list(drop_adjusted_header),
    scenario_results$result %>% map(make_original_drop_adjusted_row),
    list(drop_adjusted_footer)
  ),
  labels = c("", scenario_results$scenario, ""),
  ncol = 1,
  rel_heights = c(0.25, rep(1, nrow(scenario_results)), 0.08)
)

fig1_alt_drop_adjusted_panels

ggsave(
  filename = "figures/figure1_alternatives/fig1_alt_drop_adjusted_panels.png",
  plot = fig1_alt_drop_adjusted_panels,
  width = 150,
  height = 140,
  units = "mm",
  bg = "white"
)
```

# Alternative 2: Reviewer-Minimal

This version follows the reviewer's suggestion most directly. It removes the
four baseline-change correction panels and relies on the slope comparison for
the correction result.

```{r alternative-reviewer-minimal, fig.width=7, fig.height=7}
minimal_header <- make_column_header(
  labels = c("1. Param-\neters", "2. True\nvalues", "3. Observed\nvalues",
             "4. Slope\ncomparison"),
  rel_widths = c(0.5, 0.5, 0.5, 0.85)
)

minimal_footer <- make_column_header(
  labels = c("", "timepoint", "timepoint", "method"),
  rel_widths = c(0.5, 0.5, 0.5, 0.85),
  base_size = 8,
  face = "plain",
  vjust = -0.25
)

fig1_alt_reviewer_minimal <- cowplot::plot_grid(
  plotlist = c(
    list(minimal_header),
    scenario_results$result %>% map(make_reviewer_minimal_row),
    list(minimal_footer)
  ),
  labels = c("", scenario_results$scenario, ""),
  ncol = 1,
  rel_heights = c(0.25, rep(1, nrow(scenario_results)), 0.08)
)

fig1_alt_reviewer_minimal

ggsave(
  filename = "figures/figure1_alternatives/fig1_alt_reviewer_minimal.png",
  plot = fig1_alt_reviewer_minimal,
  width = 115,
  height = 140,
  units = "mm",
  bg = "white"
)
```

# Alternative 3: Compact Overlay

This version keeps one baseline-change panel per row. The observed points are
shown once, and fitted lines for each method are overlaid. This preserves the
visual intuition of baseline-change relationships while avoiding repeated point
clouds.

```{r alternative-compact-overlay, fig.width=8, fig.height=7}
overlay_header <- make_column_header(
  labels = c("1. Param.", "2. True", "3. Observed",
             "4. Baseline-change", "5. Slopes"),
  rel_widths = c(0.5, 0.5, 0.5, 0.78, 0.78),
  base_size = 9
)

overlay_footer <- make_column_header(
  labels = c("", "timepoint", "timepoint", "initial", "method"),
  rel_widths = c(0.5, 0.5, 0.5, 0.78, 0.78),
  base_size = 7,
  face = "plain",
  vjust = -0.25
)

fig1_alt_overlay_body <- cowplot::plot_grid(
  plotlist = c(
    list(overlay_header),
    scenario_results$result %>% map(make_overlay_row),
    list(overlay_footer)
  ),
  labels = c("", scenario_results$scenario, ""),
  ncol = 1,
  rel_heights = c(0.18, rep(1, nrow(scenario_results)), 0.08),
  label_size = 12
)

fig1_alt_overlay <- fig1_alt_overlay_body

fig1_alt_overlay

ggsave(
  filename = "figures/figure1_alternatives/fig1_alt_overlay.png",
  plot = fig1_alt_overlay,
  width = 115,
  height = 135,
  units = "mm",
  bg = "white"
)
```

# Notes

Candidate tradeoffs:

1. `fig1_alt_drop_adjusted_panels.png`: closest to the submitted layout, but
   removes the three adjusted baseline-change panels.
2. `fig1_alt_reviewer_minimal.png`: closest to the reviewer's recommendation,
   but loses the baseline-change visual argument.
3. `fig1_alt_overlay.png`: most compact while still showing how the correction
   methods alter the baseline-change slope.

If one of these becomes the manuscript Figure 1, the omitted panels from the
submitted figure can be moved into a supplemental figure generated from the
`fig1_original_recreated` object.
