---
title: "Figure 2 alternatives for reviewer response"
date: "May 2026"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE)
```

# Purpose

Reviewer feedback noted that the Blomqvist points and lines in Figure 2 are
often hidden. This notebook recreates Figure 2 and tries small graphical
adjustments before changing the analysis:

1. Put the dashed zero line behind the data.
2. Adjust transparency and draw Blomqvist last.
3. Add a small horizontal dodge if visibility remains poor.

# Setup

```{r libraries}
library(tidyverse)
library(boot)
library(future)
library(furrr)
```

```{r helpers}
dir.create("figures/figure2_alternatives", recursive = TRUE, showWarnings = FALSE)

v_pop <- 1

rttm.adj <- function(m1, m2) {
  raw.growth <- m2 - m1
  vart <- var.test(m1, m2, paired = TRUE)
  vpv <- vart$p.value
  rho <- cor.test(m1, m2)$estimate
  m1sd <- sd(m1)
  m2sd <- sd(m2)
  m1v <- var(m1)
  m2v <- var(m2)
  m1m <- mean(m1)
  m2m <- mean(m2)
  pm <- mean(raw.growth)
  rho2 <- (2 * rho * m1sd * m2sd) / (m1v + m2v)
  rhof <- ifelse(vpv <= 0.05, rho, rho2)
  dstar <- (rhof * (m1 - m1m) - (m2 - m2m)) * -1
  data.frame(raw.growth = raw.growth, dstar = dstar, adj.growth = pm + dstar)
}

moment.nohet.safe <- function(x1, x2) {
  z1 <- var(x2) - var(x1)
  z2 <- cov(x2 - x1, x1) + var(x1)
  w <- z1 / z2

  if (!is.finite(w) || abs(z2) < 1e-8) return(NA_real_)

  disc <- (2 - w)^2 + 4 * w
  if (disc < 0) return(NA_real_)

  root_plus <- (-(2 - w) + sqrt(disc)) / 2
  root_minus <- (-(2 - w) - sqrt(disc)) / 2

  b <- ifelse(cov(x1, x2) > 0, root_plus, root_minus)
  if (!is.finite(b) || abs(b) > 10) b <- NA_real_
  b
}

evaluate_methods <- function(v_pop, k, v_ind_rel, b_true, reps, n, v_err, v_ind, seed) {
  replicate_results <- purrr::map_dfr(seq_len(reps), function(i) {
    set.seed(seed + i)

    log_init_true <- rnorm(n, 0, sqrt(v_pop))
    log_change_true <- b_true * log_init_true + rnorm(n, 0, sqrt(v_ind))
    log_final_true <- log_init_true + log_change_true
    e_init <- rnorm(n, 0, sqrt(v_err))
    e_final <- rnorm(n, 0, sqrt(v_err))

    syn_df <- tibble(
      log_init_obs = log_init_true + e_init,
      log_final_obs = log_final_true + e_final,
      log_change_obs = log_final_obs - log_init_obs
    ) %>%
      mutate(log_change_kp_adj = rttm.adj(log_init_obs, log_final_obs)[, 3])

    b_obs <- coef(lm(log_change_obs ~ log_init_obs, data = syn_df))[[2]]
    b_kp <- coef(lm(log_change_kp_adj ~ log_init_obs, data = syn_df))[[2]]
    k_blomq <- v_err / (v_pop + v_err)
    b_blomq <- (b_obs + k_blomq) / (1 - k_blomq)
    b_mom <- moment.nohet.safe(syn_df$log_init_obs, syn_df$log_final_obs)
    var_p <- var.test(syn_df$log_init_obs, syn_df$log_final_obs)$p.value

    tibble(
      b_obs = b_obs,
      b_blomq = b_blomq,
      b_mom = b_mom,
      b_kp = b_kp,
      b_true = b_true,
      var_p = var_p,
      obs_bias = b_obs - b_true,
      blomq_bias = b_blomq - b_true,
      kp_bias = b_kp - b_true,
      mom_bias = b_mom - b_true
    )
  })

  replicate_results %>%
    summarise(across(
      .cols = c(b_obs, b_blomq, b_mom, b_kp, var_p, obs_bias, blomq_bias, kp_bias, mom_bias),
      .fns = list(
        mean = ~ mean(.x, na.rm = TRUE),
        sd = ~ sd(.x, na.rm = TRUE)
      ),
      .names = "{.col}_{.fn}"
    )) %>%
    mutate(
      b_true = b_true,
      v_ind = v_ind,
      v_err = v_err,
      k = k,
      v_ind_rel = v_ind_rel
    )
}

facet_labeller <- function(labels) {
  to_num <- function(x) suppressWarnings(as.numeric(as.character(x)))
  fmt <- function(x) formatC(x, format = "f", digits = 2)

  if ("k" %in% names(labels)) {
    k <- to_num(labels$k)
    k_cat <- ifelse(k <= 0.05, "Very low",
      ifelse(k <= 0.25, "Moderate",
        ifelse(k <= 0.50, "High", "Very high")
      )
    )
    labels$k <- paste0(k_cat, " (", fmt(k / (1 - k)), ")")
  }

  if ("v_ind" %in% names(labels)) {
    v <- to_num(labels$v_ind)
    v_cat <- ifelse(v <= 0.05, "Minimal",
      ifelse(v <= 1, "Substantial",
        ifelse(v <= 10, "Extreme", "Extreme")
      )
    )
    labels$v_ind <- paste0(v_cat, " (", fmt(v / v_pop), ")")
  }

  labels
}

lab_sig1 <- function(x) formatC(signif(x, 1), digits = 1, format = "fg")

method_labels <- c(
  obs_bias_mean = "Observed",
  kp_bias_mean = "Kelly-Price",
  mom_bias_mean = "Moments",
  blomq_bias_mean = "Blomqvist"
)

method_shapes <- c(
  obs_bias_mean = 16,
  kp_bias_mean = 17,
  mom_bias_mean = 15,
  blomq_bias_mean = 18
)
```

# Load Or Generate Results

The main analysis has an optional `results_grid.rds` workflow. If that file
exists, this notebook uses it. Otherwise, it regenerates the simulation grid.

```{r results-grid}
if (file.exists("results_grid.rds")) {
  results_grid <- readRDS("results_grid.rds")
} else {
  param_grid <- expand.grid(
    v_pop = 1,
    k = c(0.0099, 0.23, 0.5),
    v_ind_rel = c(5, 0.5, 0.01),
    b_true = c(-0.9, -0.5, -0.2, 0, 0.2, 0.5, 0.9),
    reps = 1000,
    n = 200
  ) %>%
    mutate(
      v_err = v_pop * k / (1 - k),
      v_ind = v_pop * v_ind_rel
    )

  set.seed(999)
  param_grid <- param_grid %>%
    mutate(seed = sample.int(1e6, size = n(), replace = FALSE))

  future::plan(multisession)
  results_grid <- furrr::future_pmap_dfr(
    param_grid,
    evaluate_methods,
    .options = furrr_options(seed = TRUE, packages = c("boot", "broom"))
  )
}

results_long_base <- results_grid %>%
  pivot_longer(
    cols = c(obs_bias_mean, kp_bias_mean, blomq_bias_mean, mom_bias_mean),
    names_to = "method",
    values_to = "error"
  ) %>%
  mutate(
    sd = case_when(
      method == "obs_bias_mean" ~ obs_bias_sd,
      method == "kp_bias_mean" ~ kp_bias_sd,
      method == "blomq_bias_mean" ~ blomq_bias_sd,
      method == "mom_bias_mean" ~ mom_bias_sd
    ),
    v_ind = factor(v_ind, levels = c(5, 0.5, 0.01))
  ) %>%
  select(v_err, k, v_ind, b_true, method, error, sd)

results_long_blomq_last <- results_long_base %>%
  mutate(method = factor(
    method,
    levels = c("obs_bias_mean", "kp_bias_mean", "mom_bias_mean", "blomq_bias_mean")
  )) %>%
  arrange(method)
```

# Plot Helper

```{r plot-helper}
plot_fig2_bias <- function(data, draw_zero_back = FALSE,
                           quiet_errorbars = FALSE,
                           point_alpha = 0.6,
                           line_alpha = 0.6,
                           dodge_width = NULL,
                           use_method_shapes = FALSE) {
  p <- ggplot(data, aes(
    x = b_true,
    y = pmax(pmin(error, 0.75), -0.75),
    color = method,
    group = method
  ))

  if (use_method_shapes) {
    p <- p + aes(shape = method)
  } else {
    p <- p + aes(shape = factor(ifelse(error > 0.75, 2, ifelse(error < -0.75, 6, 16))))
  }

  if (draw_zero_back) {
    p <- p + geom_hline(yintercept = 0, linetype = "dashed",
                        color = "black", linewidth = 0.35, alpha = 0.55)
  }

  if (is.null(dodge_width)) {
    p <- p +
      geom_line(linewidth = 0.5, alpha = line_alpha) +
      geom_errorbar(
        aes(ymin = error - sd, ymax = error + sd),
        width = 0.05,
        alpha = if (quiet_errorbars) 0.25 else 0.6
      ) +
      geom_point(size = 1.5, alpha = point_alpha)
  } else {
    dodge <- position_dodge(width = dodge_width)
    p <- p +
      geom_line(position = dodge, linewidth = 0.5, alpha = line_alpha) +
      geom_errorbar(
        aes(ymin = error - sd, ymax = error + sd),
        position = dodge,
        width = 0.04,
        alpha = if (quiet_errorbars) 0.25 else 0.6
      ) +
      geom_point(position = dodge, size = 1.5, alpha = point_alpha)
  }

  if (!draw_zero_back) {
    p <- p + geom_hline(yintercept = 0, linetype = "dashed")
  }

  p +
    scale_shape_manual(
      values = if (use_method_shapes) method_shapes else c("2" = 2, "6" = 6, "16" = 16),
      guide = if (use_method_shapes) "legend" else "none",
      breaks = if (use_method_shapes) names(method_labels) else waiver(),
      labels = if (use_method_shapes) method_labels else waiver()
    ) +
    scale_color_discrete(
      breaks = names(method_labels),
      labels = method_labels
    ) +
    facet_grid(v_ind ~ k, labeller = facet_labeller) +
    coord_cartesian(ylim = c(-0.75, 0.75)) +
    labs(
      x = "True slope of change vs. initial",
      y = "Slope error (estimate - true)",
      color = "Method",
      shape = "Method"
    ) +
    theme_minimal(base_size = 12) +
    scale_x_continuous(
      labels = lab_sig1, expand = c(0.015, 0.015),
      sec.axis = sec_axis(~ .,
        name = expression(Measurement ~ error ~~ (sigma[error]^2 / sigma[X]^2)),
        breaks = NULL, labels = NULL
      )
    ) +
    scale_y_continuous(
      sec.axis = sec_axis(~ .,
        name = expression(Change ~ heterogeneity ~~ (sigma[Delta]^2 / sigma[X]^2)),
        breaks = NULL, labels = NULL
      )
    )
}
```

# Selected Alternative: Method Shapes With Slight Dodge

This version uses method-specific point shapes, a slight horizontal dodge,
more transparent lines/points, quieter error bars, the zero reference behind
the data, and draws Blomqvist last so it sits on top.

```{r fig2-method-shapes-dodge, fig.width=6.3, fig.height=4.7}
fig2_alt_method_shapes_dodge <- plot_fig2_bias(
  results_long_blomq_last,
  draw_zero_back = TRUE,
  quiet_errorbars = TRUE,
  point_alpha = 0.55,
  line_alpha = 0.55,
  dodge_width = 0.05,
  use_method_shapes = TRUE
)

fig2_alt_method_shapes_dodge

ggsave(
  filename = "figures/figure2_alternatives/fig2_alt_method_shapes_dodge.png",
  plot = fig2_alt_method_shapes_dodge,
  width = 160,
  height = 120,
  units = "mm",
  bg = "white"
)
```
